---
title: "Bears-Breweries Study"
author: "Maryam Shahini"
date: "June 30, 2017"
output: md_document
---

# Introduction:

Todays, due to variaty of beers and brew styles, beer selection is not simple any more. Craft beers have an enormous positive impact on the economy. Craft brewing industry in the U.S. created more than 424,000 jobs and contributed $56 billion to the U.S. economy.
In this research, there are answers to some common questions regarding craft beers and breweries.  

# Import .csv data sets:
This is the first step to execute analysis. there are two .csv files received which are used in our reserach.

```{r}
Beers <- read.csv(file="C:/Users/R900255/Desktop/Uni/Doing DS/Beers.csv")
Breweries <- read.csv(file="C:/Users/R900255/Desktop/Uni/Doing DS/Breweries.csv")

```

Before starting any research, it is important to understand available sources, columns and information. and clean data sets if needed.
Beers dataset contains a list of 2410 US craft beers and 7 columns of data and Breweries dataset contains 558 US
breweries and 4 columns of data. 
The datasets descriptions and columns included are as follows:

Beers.csv:
Name: Name of the beer.
Beer ID: Unique identi er of the beer.
ABV: Alcohol by volume of the beer.
IBU: International Bitterness Units of the beer.
Brewery ID: Brewery id associated with the beer.
Style: Style of the beer.
Ounces: Ounces of beer.

Breweries.csv:
Brew ID: Unique identi er of the brewery.
Name: Name of the brewery.
City: City where the brewery is located.
State: State where the brewery is located.

```{r}
names(Beers)
dim(Beers)
names(Breweries)
dim(Breweries)
```


# 1. The number of breweries present in each state:

This study shows that the number of breweries present in different states are different.
CO has the maximum number of breweries.

1       AK  7
2       AL  3
3       AR  2
4       AZ 11
5       CA 39
6       CO 47
7       CT  8
8       DC  1
9       DE  2
10      FL 15
11      GA  7
12      HI  4
13      IA  5
14      ID  5
15      IL 18
16      IN 22
17      KS  3
18      KY  4
19      LA  5
20      MA 23
21      MD  7
22      ME  9
23      MI 32
24      MN 12
25      MO  9
26      MS  2
27      MT  9
28      NC 19
29      ND  1
30      NE  5
31      NH  3
32      NJ  3
33      NM  4
34      NV  2
35      NY 16
36      OH 15
37      OK  6
38      OR 29
39      PA 25
40      RI  5
41      SC  4
42      SD  1
43      TN  3
44      TX 28
45      UT  4
46      VA 16
47      VT 10
48      WA 23
49      WI 20
50      WV  1
51      WY  4

```{r}
Brew.per.State  <- aggregate(rep(1, length(paste0(Breweries$State))),
                             by=list(Breweries$State), sum)
Brew.per.State
```

#2. Merge beer data with breweries data by brewery id:

Print frst 6 observations and the last six observations to check the merged file.

In order to study available data, first checking the column names to make sure that they are clear in meaning and to understand the information included in each table.
A key column in each table used to joint tables.

Key column is "Brewery_id" in Beers data set and "Brew_ID"" in Breweries data set.
In order to have the same name for key columns in both data sets, it needs to rename the "Brew_ID" column to "Brewery_id".

After renaming "Brew_ID" column, the list of columns in "Breweries" dataset update to:
"Brewery_id"
"Name"
"City"
"State" 

```{r}
colnames(Breweries)[colnames(Breweries)=="Brew_ID"]<-"Brewery_id"
names(Breweries)
names(Beers)
```

Now, it is time to merge data sets. A new object "BeerBrew" which is the result of merging tables is created. 
The dimension of the "BeerBrew" data frame is 2410 rows and 10 columns (combination of Beers columns and Breweries columns). columns are as below:
"Brewery_id"
"Name.Beers"
"Beer_ID"
"ABV"  
"IBU"  
"Style"
"Ounces" 
"Name.Breweries"
"City"
"State"

in the merged file, top 6 rows are as below:

Row  Brewery_id   Name.Beers Beer_ID  ABV IBU                      Style Ounces     Name.Breweries          City       State
1          1  Get Together    2692 0.045  50                        American IPA     16 NorthGate Brewing  Minneapolis    MN
2          1 Maggie's Leap    2691 0.049  26                  Milk / Sweet Stout     16 NorthGate Brewing  Minneapolis    MN
3          1    Wall's End    2690 0.048  19                   English Brown Ale     16 NorthGate Brewing  Minneapolis    MN
4          1       Pumpion    2689 0.060  38                         Pumpkin Ale     16 NorthGate Brewing  Minneapolis    MN
5          1    Stronghold    2688 0.060  25                     American Porter     16 NorthGate Brewing  Minneapolis    MN
6          1   Parapet ESB    2687 0.056  47 Extra Special / Strong Bitter (ESB)     16 NorthGate Brewing  Minneapolis    MN

and last 6 rows are as below:

Row  Brewery_id           Name.Beers       Beer_ID   ABV IBU         Style           Ounces          Name.Breweries          City      State
2405        556             Pilsner Ukiah      98 0.055  NA         German Pilsener     12         Ukiah Brewing Company         Ukiah    CA
2406        557  Heinnieweisse Weissebier      52 0.049  NA              Hefeweizen     12       Butternuts Beer and Ale Garrattsville    NY
2407        557           Snapperhead IPA      51 0.068  NA            American IPA     12       Butternuts Beer and Ale Garrattsville    NY
2408        557         Moo Thunder Stout      50 0.049  NA      Milk / Sweet Stout     12       Butternuts Beer and Ale Garrattsville    NY
2409        557         Porkslap Pale Ale      49 0.043  NA American Pale Ale (APA)     12       Butternuts Beer and Ale Garrattsville    NY
2410        558 Urban Wilderness Pale Ale      30 0.049  NA        English Pale Ale     12 Sleeping Lady Brewing Company     Anchorage    AK

```{r}
BeerBrew<- merge(x=Beers,y=Breweries, by= "Brewery_id",
                       all.x = TRUE, sort = TRUE, suffixes = c(".Beers", ".Breweries"),
                       allow.cartesian=getOption("datatable.allow.cartesian"))
dim(BeerBrew)
names(BeerBrew)
head(BeerBrew,6)
tail(BeerBrew,6)

```

# 3. The number of NA's in each column.

Both Beers and Breweries data sets may have NA values in some of their columns. the number of NA's in each columns are shown below:

In Beers data set, ABV column has 62 NA's and IBU column has 1005 NA's:
Name    Beer_ID        ABV        IBU    Brewery_id      Style     Ounces 
 0          0          62        1005          0          0          0 

In Breweries data set, none of the columns has NA's:
 Brewery_id     Name       City      State 
      0          0          0          0 
In BeerBrew file, ABV column has 62 NA's and IBU column has 1005 NA's:
Brewery_id     Name.Beers        Beer_ID        ABV            IBU          Style         Ounces    Name.Breweries        City      State
     0              0              0             62           1005              0              0            0              0          0
        
```{r}
colSums(is.na(Beers))
colSums(is.na(Breweries))
colSums(is.na(BeerBrew))
```

# 4. The median alcohol content and international bitterness unit for each state:

First, two new files which are subset of BeerBrew data frame are created. "IBUclean" and "ABVclean" do not include missing values of IBU and ABV. it is required to ignore missing values for calculating medians.
ABVclean data frame includes 2348 rows and 10 columns.
IBUclean data frame includes 1405 rows and 10 columns.

```{r}
IBUclean<-subset(x=BeerBrew, ! is.na(IBU))
ABVclean<-subset(x=BeerBrew, ! is.na(ABV))
dim(ABVclean)
dim(IBUclean)
```

The method used to calculate the medians of ABV and IBU per state in this study required to install "doBy" package. 
IBUmedian file includes IBU median per state by using IBUclean data frame.
ABVmedian file also includes ABV median per state by using ABVclean data frame.
Then both IBUmedian and ABVmedian files are merged to have calculated medians of ABV and IBU in the same file "New".

the median of ABV and IBU in each state is as below:
   State ABV.median IBU.median
1     AK     0.0560       46.0
2     AL     0.0600       43.0
3     AR     0.0520       39.0
4     AZ     0.0550       20.5
5     CA     0.0580       42.0
6     CO     0.0605       40.0
7     CT     0.0600       29.0
8     DC     0.0625       47.5
9     DE     0.0550       52.0
10    FL     0.0570       55.0
11    GA     0.0550       55.0
12    HI     0.0540       22.5
13    IA     0.0555       26.0
14    ID     0.0565       39.0
15    IL     0.0580       30.0
16    IN     0.0580       33.0
17    KS     0.0500       20.0
18    KY     0.0625       31.5
19    LA     0.0520       31.5
20    MA     0.0540       35.0
21    MD     0.0580       29.0
22    ME     0.0510       61.0
23    MI     0.0620       35.0
24    MN     0.0560       44.5
25    MO     0.0520       24.0
26    MS     0.0580       45.0
27    MT     0.0550       40.0
28    NC     0.0570       33.5
29    ND     0.0500       32.0
30    NE     0.0560       35.0
31    NH     0.0550       48.5
32    NJ     0.0460       34.5
33    NM     0.0620       51.0
34    NV     0.0600       41.0
35    NY     0.0550       47.0
36    OH     0.0580       40.0
37    OK     0.0600       35.0
38    OR     0.0560       40.0
39    PA     0.0570       30.0
40    RI     0.0550       24.0
41    SC     0.0550       30.0
42    SD     0.0600         NA
43    TN     0.0570       37.0
44    TX     0.0550       33.0
45    UT     0.0400       34.0
46    VA     0.0565       42.0
47    VT     0.0550       30.0
48    WA     0.0555       38.0
49    WI     0.0520       19.0
50    WV     0.0620       57.5
51    WY     0.0500       21.0

```{r}
library(doBy)
IBUmedian<-summaryBy(IBU~State, data=IBUclean,FUN=list(median), na.rm=F)
ABVmedian<-summaryBy(ABV~State, data=ABVclean,FUN=list(median), na.rm=F)
New<-merge(x=ABVmedian, y=IBUmedian, by="State", all=TRUE)
dim(New)
names(New)
New
```

# Bar chart of median of ABV and IBU for each state:

The bar chart is created by using "ggplot2" package and "new" file. The first bar chart shows the ABV median versus IBU median and each bar represent one state. 
the next bar chart is the ABV median versus state. The last one is IBU median versus state.
As shown in different bar charts, DC and KY have the maximum ABV median and ME has the maximum IBU median among all states.

```{r}

library(ggplot2)
ggplot(data=New, aes(x=IBU.median, y=ABV.median, fill=State))+geom_bar(stat="identity",position=position_dodge())+theme(axis.text.x=element_text(angle=90,vjust=0.5))
ggplot(data=New, aes(x=State, y=ABV.median, fill=State))+geom_bar(stat="identity",position=position_dodge())+theme(axis.text.x=element_text(angle=90,vjust=0.5))
ggplot(data=New, aes(x=State, y=IBU.median, fill=State))+geom_bar(stat="identity",position=position_dodge())+theme(axis.text.x=element_text(angle=90,vjust=0.5))
```

# 5. Which state has the maximum alcoholic beer? Which state has the most bitter beer?

BeerBrew data frame (the merge of Beers and Breweries data sets) is used to find states with maximum ABV and IBU, as BeerBrew includes all state column and IBU and ABV columns.
CO has the maximum ABV of 0.128 and OR has the maximum IBU of 138.

```{r}
BeerBrew[which.max(BeerBrew$ABV),]
BeerBrew[which.max(BeerBrew$IBU),]
```

# 6. Summary statistics for ABV (Alcohol by volume) variable.

Basic statistics of ABV variable is calculated from "ABVclean" table as the missing values of ABV is ignored in this table. 
Minimum = 0.00100  /   1st Qu. = 0.05000  /  median = 0.05600  /  mean = 0.05977  /  3rd Qu. = 0.06700  and maximum = 0.128

```{r}
summary(ABVclean$ABV)
```

# 7. Relationship between the bitterness of the beer and its alcoholic content

By drawing a scatter plot of ABV versus IBU, it is obvious that there is a positive correlation between ABV and IBU. The fit line has a positive slope that indicates that with increasing ABV variable, IBU also increased. that meand beers with higher alcoholic content tend to have higher bitterness.

```{r}
plot(x=Beers$ABV,y=Beers$IBU,main="Compare",xlab="ABV", ylab="IBU")
abline(lm(Beers$IBU~Beers$ABV), col="red")
```


# 8. Conclusion:

This study is a reproducible research that answers some commeon questions by using craft beers and breweries information. By using methods done in this study, we are able to perform many more analysis and find reliable answers for other questions.
As shown in abov descriptions, number of breweries in each state, number of NA's in each column, IBU and ABV median calculation for each state, states with maximum alcoholic beer and the most bitter beer, summary statistic of ABV variable and relationships between the bitterness and alcoholic content of beers are presented. There are also a few bar charts and scatter plot that clarify some of our results. 




